Supplementary Information -- Morphological convergence as on-line lexical analogy

### 2020-01-30 supplementary_information/si.Rnw
for clay

# libs
library(tidyverse)
library(xtable)
library(lme4)
library(patchwork) # take this out

@@

# setting up columns in baseline and post-test data
baseline = baseline %>% 
  mutate(
    r_baseline_mgl_features = scales::rescale(baseline_mgl_features), # rescale regularity weights
    r_baseline_gcm_features = scales::rescale(baseline_gcm_features),
  )

test2 = test2 %>% 
  mutate(
    verb_baseline_average = baseline_av - mean(baseline_av, na.rm = T), # center verb average
    participant_prereg_average = pre_reg_av - mean(pre_reg_av, na.rm = T), # center participant average
    bot_lexical_typicality = case_when( # set factor levels used in paper, order factor
        lex_typicality == 'typical' ~ 'typical',
        lex_typicality == 'random' ~ 'random',
        lex_typicality == 'atypical' ~ 'reversed'
      ) %>% 
      factor(levels = c('typical', 'random', 'reversed')),
    bot_regularisation_rate = reg_rate %>% # order factor
      factor(levels = c('nc', '-40%', '+40%')),
    r_baseline_mgl_features = scales::rescale(baseline_mgl_features), # scale regular weights
    r_baseline_gcm_features = scales::rescale(baseline_gcm_features),
    r_individual_mgl_features = scales::rescale(individual_mgl_features),
    r_individual_gcm_features = scales::rescale(individual_gcm_features),
    gcm_score_diff = r_individual_gcm_features - r_baseline_gcm_features, # calculate score diff
    mgl_score_diff = r_individual_mgl_features - r_baseline_mgl_features
  )

@@

## get the raw output of the baseline feature-distance mgl
mgl_baseline_rules_raw = read_tsv('~/Github/RaczBecknerHayPierrehumbert2019/models/mgl/baseline_mgl/CELEXFull3.sum') %>% 
  mutate(
    form1 = str_replace_all(form1, '»', ''), # get rid of syllable boundary markers
    form2 = str_replace_all(form2, '»', '')
  )

## get verb handles
verb_handles = read_csv('~/Github/RaczBecknerHayPierrehumbert2019/data/256verbs_print_forms.txt')

@@

We also discuss (\ref{appendixgcmmglresidualization}) the residualization method used to compare them, (\ref{posttestimplementation}) how well they fit our ESP test, and (\ref{appendixregression}) the regression methods used.

@@

# table 1
baseline %>% 
  group_by(word) %>% 
  summarise(mean = mean(regular)) %>% 
  arrange(-mean)

@@

gcm_table = celex %>% # take celex
  filter(!is.na(ahpa)) %>% # remove words not in Albright and Hayes celex
  mutate(
    vowels = str_extract_all(ahpa, '[euVoEIi2U{Q34@a6}]'), # match vowels in the albright and hayes transcription
    nvowels = map(vowels, ~ length(.)) # calculate how many vowels the verb has
  ) %>% 
  select(-vowels) %>% # we don't need vowels now
  unnest(cols = c(nvowels)) %>%  # unnest nvowels
  filter(word != 'forsake', nvowels == 1) %>% # drop any verb with more than 1 vowel also "forsake which survives this
  select(category, moder, disc) %>% # keep category, moder, disc columns (= reg/irreg, drove/sang etc, transcription)
  distinct() %>% # repetitions! 
  count(category, moder) %>% # count drove/sang etc across reg/irreg
  mutate( # if it has no category, it's a misc verb
    moder = replace_na(moder, 'other')
  ) %>% 
  pivot_wider( # stretch table
    names_from = category, values_from = n
  ) %>% 
  rename(
   `verb class` = moder,
    `irregular set` = irregular,
    `regular set` = regular
  )

reg_count = filter(gcm_table, `verb class` == 'other') %>% 
  pull(`regular set`)

gcm_table = gcm_table[1:4,] %>% 
  mutate(
    `miscellaneous regular verbs` = reg_count
  )

@@

<<gcmtable, echo=F, eval=T, results=tex>>=
print(xtable(gcm_table, digits=2, label="gcmsetsize", caption="Number of forms in each verb class, GCM training data"))

@@

irreg_set = gcm_table %>% 
  filter(`verb class` == 'DROVE') %>% 
  pull(`irregular set`) %>% 
  as.double

reg_set = gcm_table %>% 
  filter(`verb class` == 'DROVE') %>% 
  pull(`regular set`) %>% 
  as.double

splive_gcm = test2 %>% 
  filter(word == 'splive') %>% 
  pull(r_baseline_gcm_features) %>% 
  unique %>% 
  round(2)

@@

%The nonce form {\em splive} belongs to the DROVE class in our model. The two past forms of {\em splive} in the experiment are regular {\em splived} and irregular {\em splove}. It is compared to \Sexpr{reg_set + reg_count} regular verbs -- these are \Sexpr{reg_set} verbs that match the DROVE schema (e.g.\ {\em side, hive, line}) and \Sexpr{reg_count} miscellaneous verbs. It is also compared to \Sexpr{irreg_set} irregular verbs (e.g.\ {\em drive, stride, smite}) in this class. Overall, it is more similar to the regular set: its {\em regularity score} is \Sexpr{splive_gcm}.

@@

disc_category = test2 %>% # get words and categories from test2. match colnames with above
  distinct(disc, category) %>% 
  mutate(
    form1 = disc
  )

disc_reg_past_irreg_past = verb_handles %>% # get reg and irreg transcriptions from verb handles. match colnames with above.
  select(base.form,regular.form,irregular.form) %>% 
  rename(
    form1 = base.form
  )

@@

mgl_baseline_rules_raw = mgl_baseline_rules_raw %>% 
  filter(
    !is.na(category)
  )

@@

mgl_baseline_rules_raw = mgl_baseline_rules_raw %>%  # group rules per output: reg / irreg!
  mutate(
    rule.type = case_when(
      form2 == regular.form ~ 'regular',
      form2 == irregular.form ~ 'irregular'
    )
  )

mgl_baseline_rules_raw = mgl_baseline_rules_raw %>%
  filter(
!is.na(rule.type)
)

@@

mgl_count = mgl_baseline_rules_raw %>% # clean-up
  mutate(
    `related forms` = str_replace_all(`related forms`, '»', ''),
    exceptions = str_replace_all(exceptions, '»', ''),
    related_list = strsplit(`related forms`, ', '), # split
    exception_list = strsplit(exceptions, ', ')
  )

mgl_count1 = mgl_count %>%
  select(category,rule.type,related_list) %>% 
  unnest(related_list) %>% 
  distinct() %>% 
  count(category, rule.type) %>% 
  mutate(
    type = 'related forms'
  )

mgl_count2 = mgl_count %>%
  select(category,rule.type,exception_list) %>% 
  unnest(exception_list) %>% 
  distinct() %>% 
  count(category, rule.type) %>% 
  mutate(
    type = 'exceptions'
  )

mgl_table = rbind(mgl_count1,mgl_count2) %>% 
  pivot_wider(
    names_from = type, values_from = n
  ) %>% 
  arrange(rule.type, category)

@@

<<mgltable, echo=F, eval=T, results=tex>>=
print(xtable(mgl_table, digits=2, label="mglsetsize", caption="Number of forms in each verb class, MGL training data"))

@@

# In the next par, I won't pipe the values into variables in the text, but you can look them up here:
splive_mgl_rules = mgl_baseline_rules_raw %>% 
  filter(
    form1 == 'spl2v'
  )
# regular
splive_mgl_rules[3,]$Pfeat
splive_mgl_rules[3,]$hits
splive_mgl_rules[3,]$scope
splive_mgl_rules[3,]$reliability
splive_mgl_rules[3,]$confidence

@@

test2 %>% 
  mutate(
    r_baseline_mgl_features = scales::rescale(baseline_mgl_features)
  ) %>% 
  filter(word == 'splive') %>% 
  select(baseline_mgl_features,r_baseline_mgl_features) %>% 
  distinct()

@@

mgl_rule_count = mgl_baseline_rules_raw %>% # take raw rules
  filter( # throw out non-category verbs and non reg/irreg rules (in our definition)
    !is.na(category),
    !is.na(rule.type)
  ) %>% 
  mutate( # make a rule column of the form A -> B / C
    rule = paste0(A, ' ->', B, ' / ', Pfeat)
  ) %>% 
  distinct(rule,category,rule.type) %>% # keep DISTINCT rules
  count(category, rule.type) %>%  # count them
  pivot_wider(names_from = rule.type, values_from = n) %>% # stretch table
  mutate( # make row sum
    sum = irregular + regular
  ) %>% 
  rename(
    `regular rule` = regular,
    `irregular rule` = irregular
  )

# glue on colsums
mgl_rule_count = rbind(mgl_rule_count, c('sum', sum(mgl_rule_count$`irregular rule`), sum(mgl_rule_count$`regular rule`), sum(mgl_rule_count$sum)))

@@

<<mgltable, echo=F, eval=T, results=tex>>=
print(xtable(mgl_table, digits=2, label="mglruleses", caption="The number of unique rules posited by the MGL in each verb class"))

@@

\begin{table}[ht]
\centering
\begin{tabular}{lll}
  \hline
alternation & preceding environment\\ 
  \hline
\textipa{il} $\rightarrow{}$ \textipa{Elt} & obstruent \\ 
\textipa{ip} $\rightarrow{}$ \textipa{Ept} & \{j, l, r, w\} \\ 
\textipa{ip} $\rightarrow{}$ \textipa{Ept} & dorsal consonant\\
\textipa{ip} $\rightarrow{}$ \textipa{Ept} & l \\ 
\textipa{ip} $\rightarrow{}$ \textipa{Ept} & any consonant \\
\textipa{il} $\rightarrow{}$ \textipa{Elt} & fricative or nasal consonant \\
$\emptyset{}$ $\rightarrow{}$ \textipa{d}    & alveolars \\ 
$\emptyset{}$ $\rightarrow{}$ \textipa{d}    & \{b, m\} \\ 
$\emptyset{}$ $\rightarrow{}$ \textipa{d}    & alveolar fricative or nasal consonant \\ 
$\emptyset{}$ $\rightarrow{}$ \textipa{d}    & \{b, m, p\} \\
   \hline
\end{tabular}
\caption{The unique rules posited for the KEPT class}
\label{mglruleseskept}
\end{table}

@@

##########################################
# from Florian Jaeger

@@

@@

save(lme.fit1, file = 'si_fit1.rda')
save(lme.fit1b, file = 'si_fit1b.rda')
save(lme.fit1c, file = 'si_fit1c.rda')
vif.mer(lme.fit1)
vif.mer(lme.fit1c)

@@

@@

<<main_model3, echo=F,eval=T, results=tex>>=
load('si_fit1.rda')
print(xtable(summary(lme.fit1)$coef, digits = 3, label = 'regressiontable1', caption = 'Experimental Data, Post-test Regression Model Summary. Model formula: post-test regular response ~ verb baseline mean x lexical typicality + participant pretest mean + regularization shift + (1 + verb.baseline.mean | participant) + (1 | verb)'))
@

@@

<<Bayes factor, echo=F, eval=F>>=
# random vs atypical

test2b = test2 %>% 
  filter(
    bot_lexical_typicality != 'typical'
  )

@@

# summary(fit_power1)
# summary(fit_power2)
save(fit_power1, file = 'si_fit_power1.rda')
save(fit_power2, file = 'si_fit_power2.rda')

exp((BIC(fit_power2) - BIC(fit_power1))/2)
@

@@

@@

mgl_gcm_cor = baseline %>% 
  distinct(word,r_baseline_mgl_features,baseline_gcm_features) %>% 
  summarise(
    cor = cor(r_baseline_mgl_features,baseline_gcm_features)
  ) %>% 
  pull() %>% 
  round(3)

@@

baseline_words = baseline %>% 
  distinct(word, r_baseline_mgl_features, r_baseline_gcm_features)

celex_m1 = lm(r_baseline_gcm_features ~ 1 + r_baseline_mgl_features, data = baseline_words)
celex_m2 = lm(r_baseline_mgl_features ~ 1 + r_baseline_gcm_features, data = baseline_words)

baseline_words = baseline_words %>% 
  mutate(
    r_baseline_gcm_features_residual = resid(celex_m1),
    r_baseline_mgl_features_residual = resid(celex_m2)
  )

baseline = left_join(baseline, baseline_words, by = c("word", "r_baseline_mgl_features", "r_baseline_gcm_features"))

celex_m3 = glmer(regular ~ 
                1 + r_baseline_mgl_features + r_baseline_gcm_features_residual + 
                (1 + r_baseline_mgl_features + r_baseline_gcm_features_residual | subject), 
                data = baseline, 
                family = binomial ( link = logit ),
                control=glmerControl(optimizer="bobyqa"))

@@

save(celex_m3, file = 'si_celex_m3.rda')
save(celex_m4, file = 'si_celex_m4.rda')

@@

<<celex_resid1, echo=F,eval=T, results=tex>>=
load('si_celex_m3.rda')
print(xtable(summary(celex_m3)$coef, digits = 3, label = 'resid1', caption = 'MGL scores and GCM residuals, baseline data'))
@

<<celex_resid2, echo=F,eval=T, results=tex>>=
load('si_celex_m4.rda')
print(xtable(summary(celex_m4)$coef, digits = 3, label = 'resid2', caption = 'GCM scores and MGL residuals, baseline data'))
@

@@

\section{Model fits on the post-test data}

@@

@@

@@

save(indiv_m5, file = 'si_indiv_m5.rda')

@@

save(indiv_m6, file = 'si_indiv_m6.rda')

@@

load('si_indiv_m5.rda')
print(xtable(summary(indiv_m5)$coef, digits = 3, label = 'indiv1', caption = 'Celex-GCM and individual score difference, post-test data'))

@@

<<indiv_diff2, echo=F,eval=T, results=tex>>=
load('si_indiv_m6.rda')
print(xtable(summary(indiv_m6)$coef, digits = 3, label = 'indiv2', caption = 'Celex-MGL and individual score difference, post-test data'))
@

@@

save(comp_fit1, file = 'si_comp_fit1.rda')
save(comp_fit1b, file = 'si_comp_fit1b.rda')
save(comp_fit1c, file = 'si_comp_fit1c.rda')
save(comp_fit1d, file = 'si_comp_fit1d.rda')
save(comp_fit1e, file = 'si_comp_fit1e.rda')

@@

load('si_comp_fit1.rda')
load('si_comp_fit1b.rda')
load('si_comp_fit1c.rda')
load('si_comp_fit1d.rda')
load('si_comp_fit1e.rda')

@@

<<comparison_model3, echo=F,eval=T, results=tex>>=
print(xtable(summary(comp_fit1)$coef, digits = 3, label = 'comp3', caption = 'Celex-MGL and Individual-GCM predict variation in the post-test'))
@

@@

# can't figure first panel out

first_panel = test2 %>% 
  group_by(
    word,
    bot_lexical_typicality,
    baseline_av
  ) %>%
  summarise(
    post_reg_mean = post_reg_av %>% 
      mean()
  ) %>% 
  ungroup() %>% 
  mutate(
    baseline = scales::rescale(baseline_av),
    posttest = scales::rescale(post_reg_mean),
    type = 'Participant responses'
  )

second_panel = test2 %>% 
  group_by(
    word,
    bot_lexical_typicality,
    r_baseline_gcm_features
  ) %>% 
  rename(
    baseline = r_baseline_gcm_features
  ) %>% 
  summarise(
    posttest = mean(r_individual_gcm_features)
  ) %>% 
  mutate(
    type = 'GCM'
  )

third_panel = test2 %>% 
  group_by(
    word,
    bot_lexical_typicality,
    r_baseline_mgl_features
  ) %>% 
  rename(
    baseline = r_baseline_mgl_features
  ) %>% 
  summarise(
    posttest = mean(r_individual_mgl_features)
  ) %>% 
  mutate(
    type = 'MGL'
  )

# three_studies = bind_rows(first_panel,second_panel,third_panel)  
three_studies = bind_rows(second_panel,third_panel)

p1 = ggplot(three_studies, aes(x = baseline, y = posttest, colour = bot_lexical_typicality, shape = bot_lexical_typicality, linetype = bot_lexical_typicality)) +
  guides(size = F) +
  geom_point() +
  geom_smooth(se = F, method = 'loess', formula = 'y ~ x') +
  facet_wrap( ~ type) +
  scale_shape_discrete(name="Lexical\ntypicality") +
  # scale_colour_discrete(name="Lexical\ntypicality") +
  xlab('verb baseline regularization') +
  ylab('verb ESP regularization') +
  guides(linetype = F) +
  scale_colour_brewer(palette = 'Set2', name="Lexical\ntypicality")

>>>>

# this SI uses tidy data manipulation -- for a quick primer, visit https://dplyr.tidyverse.org/

# libraries
library(tidyverse) # for data wrangling, figures
library(purrr) # for functional programming
library(stringr) # for string manipulation
library(readr) # for reading
library(scales) # for rescaling
library(xtable) # for latex tables
library(lme4) # for GLMM-s

@@

# setting up columns in baseline and post-test data
baseline = baseline %>% 
  dplyr::mutate(
    r_baseline_mgl_features = scales::rescale(baseline_mgl_features), # rescale regularity weights
    r_baseline_gcm_features = scales::rescale(baseline_gcm_features),
  )

test2 = test2 %>% 
  dplyr::mutate(
    verb_baseline_average = baseline_av - mean(baseline_av, na.rm = T), # center verb average
    participant_prereg_average = pre_reg_av - mean(pre_reg_av, na.rm = T), # center participant average
    bot_lexical_typicality = dplyr::case_when( # set factor levels used in paper, order factor
        lex_typicality == 'typical' ~ 'typical',
        lex_typicality == 'random' ~ 'random',
        lex_typicality == 'atypical' ~ 'reversed'
      ) %>% 
      factor(levels = c('typical', 'random', 'reversed')),
    bot_regularisation_rate = reg_rate %>% # order factor
      factor(levels = c('nc', '-40%', '+40%')),
    r_baseline_mgl_features = scales::rescale(baseline_mgl_features), # scale regular weights
    r_baseline_gcm_features = scales::rescale(baseline_gcm_features),
    r_individual_mgl_features = scales::rescale(individual_mgl_features),
    r_individual_gcm_features = scales::rescale(individual_gcm_features),
    gcm_score_diff = r_individual_gcm_features - r_baseline_gcm_features, # calculate score diff
    mgl_score_diff = r_individual_mgl_features - r_baseline_mgl_features
  )

@@

## get the raw output of the baseline feature-distance mgl
mgl_baseline_rules_raw = readr::read_tsv('~/Github/RaczBecknerHayPierrehumbert2019/models/mgl/baseline_mgl/CELEXFull3.sum') %>% 
  dplyr::mutate(
    form1 = stringr::str_replace_all(form1, '»', ''), # get rid of syllable boundary markers
    form2 = stringr::str_replace_all(form2, '»', '')
  )

## get verb handles
verb_handles = readr::read_csv('~/Github/RaczBecknerHayPierrehumbert2019/data/256verbs_print_forms.txt')

@@

We also discuss model selection in analysing our regression data (\ref{appendixregression}), how the two learning models compare on our baseline data (\ref{appendixgcmmglresidualization}), and how they compare on our ESP post-test data (\ref{mglgcmposttest}).

@@

# table 1
baseline %>% 
  dplyr::group_by(word) %>% 
  dplyr::summarise(mean = mean(regular)) %>% 
  dplyr::arrange(-mean)

@@

gcm_table = celex %>% # take celex
  dplyr::filter(!is.na(ahpa)) %>% # remove words not in Albright and Hayes celex
  dplyr::mutate(
    vowels = stringr::str_extract_all(ahpa, '[euVoEIi2U{Q34@a6}]'), # match vowels in the albright and hayes transcription
    nvowels = purrr::map(vowels, ~ length(.)) # calculate how many vowels the verb has
  ) %>% 
  dplyr::select(-vowels) %>% # we don't need vowels now
  tidyr::unnest(cols = c(nvowels)) %>%  # unnest nvowels
  dplyr::filter(word != 'forsake', nvowels == 1) %>% # drop any verb with more than 1 vowel also "forsake which survives this
  dplyr::select(category, moder, disc) %>% # keep category, moder, disc columns (= reg/irreg, drove/sang etc, transcription)
  dplyr::distinct() %>% # repetitions! 
  dplyr::count(category, moder) %>% # count drove/sang etc across reg/irreg
  dplyr::mutate( # if it has no category, it's a misc verb
    moder = tidyr::replace_na(moder, 'other')
  ) %>% 
  tidyr::pivot_wider( # stretch table
    names_from = category, values_from = n
  ) %>% 
  dplyr::rename(
   `verb class` = moder,
    `irregular set` = irregular,
    `regular set` = regular
  )

reg_count = dplyr::filter(gcm_table, `verb class` == 'other') %>% 
  dplyr::pull(`regular set`)

gcm_table = gcm_table[1:4,] %>% 
  dplyr::mutate(
    `miscellaneous regular verbs` = reg_count
  )

@@

<<gcmtable, echo=F, eval=T, results=tex>>=
print(xtable::xtable(gcm_table, digits=2, label="gcmsetsize", caption="Number of forms in each verb class, GCM training data"))

@@

irreg_set = gcm_table %>% 
  dplyr::filter(`verb class` == 'DROVE') %>% 
  dplyr::pull(`irregular set`) %>% 
  as.double

reg_set = gcm_table %>% 
  dplyr::filter(`verb class` == 'DROVE') %>% 
  dplyr::pull(`regular set`) %>% 
  as.double

splive_gcm = test2 %>% 
  dplyr::filter(word == 'splive') %>% 
  dplyr::pull(r_baseline_gcm_features) %>% 
  unique %>% 
  round(2)

@@

The nonce form {\em splive} belongs to the DROVE class in our model. The two past forms of {\em splive} in the experiment are regular {\em splived} and irregular {\em splove}. It is compared to \Sexpr{reg_set + reg_count} regular verbs -- these are \Sexpr{reg_set} verbs that match the DROVE schema (e.g.\ {\em side, hive, line}) and \Sexpr{reg_count} miscellaneous verbs. It is also compared to \Sexpr{irreg_set} irregular verbs (e.g.\ {\em drive, stride, smite}) in this class. Overall, it is more similar to the regular set: its {\em regularity score} is \Sexpr{splive_gcm}.

@@

disc_category = test2 %>% # get words and categories from test2. match colnames with above
  dplyr::distinct(disc, category) %>% 
  dplyr::mutate(
    form1 = disc
  )

disc_reg_past_irreg_past = verb_handles %>% # get reg and irreg transcriptions from verb handles. match colnames with above.
  dplyr::select(base.form,regular.form,irregular.form) %>% 
  dplyr::rename(
    form1 = base.form
  )

@@

mgl_baseline_rules_raw = mgl_baseline_rules_raw %>% 
  dplyr::filter(
    !is.na(category)
  )

@@

mgl_baseline_rules_raw = mgl_baseline_rules_raw %>%  # group rules per output: reg / irreg!
  dplyr::mutate(
    rule.type = dplyr::case_when(
      form2 == regular.form ~ 'regular',
      form2 == irregular.form ~ 'irregular'
    )
  )

mgl_baseline_rules_raw = mgl_baseline_rules_raw %>%
  dplyr::filter(
!is.na(rule.type)
)

@@

mgl_count = mgl_baseline_rules_raw %>% # clean-up
  dplyr::mutate(
    `related forms` = stringr::str_replace_all(`related forms`, '»', ''),
    exceptions = stringr::str_replace_all(exceptions, '»', ''),
    related_list = strsplit(`related forms`, ', '), # split related forms list, a long string, into a character vector, store the vector
    exception_list = strsplit(exceptions, ', ') # same for exceptions
  )

mgl_count1 = mgl_count %>%
  dplyr::select(category,rule.type,related_list) %>% 
  tidyr::unnest(related_list) %>% # we now have the related forms as one verb per line, which is tidy. but the verbs repeat, since the same verb can be a related form for several target verbs
  dplyr::distinct() %>% # so we distinct them!
  dplyr::count(category, rule.type) %>% # since one distinct verb related form == one row, we can count the rows
  dplyr::mutate(
    type = 'related forms' # by the way, this is related forms
  )

mgl_count2 = mgl_count %>% # we repeat this for exceptions
  dplyr::select(category,rule.type,exception_list) %>% 
  tidyr::unnest(exception_list) %>% 
  dplyr::distinct() %>% 
  dplyr::count(category, rule.type) %>% 
  dplyr::mutate(
    type = 'exceptions'
  )

mgl_table = rbind(mgl_count1,mgl_count2) %>%  # we combine the two
  tidyr::pivot_wider( # stretch the data frame
    names_from = type, values_from = n
  ) %>% 
  dplyr::arrange(rule.type, category)

@@

<<mgltable, echo=F, eval=T, results=tex>>=
print(xtable::xtable(mgl_table, digits=2, label="mglsetsize", caption="Number of forms in each verb class, MGL training data"))

@@

# In the next par, I won't pipe the values into variables in the text, but you can look them up here:
splive_mgl_rules = mgl_baseline_rules_raw %>% 
  dplyr::filter(
    form1 == 'spl2v'
  )
# regular
splive_mgl_rules[3,]$Pfeat
splive_mgl_rules[3,]$hits
splive_mgl_rules[3,]$scope
splive_mgl_rules[3,]$reliability
splive_mgl_rules[3,]$confidence

@@

test2 %>% 
  dplyr::filter(word == 'splive') %>% 
  dplyr::select(baseline_mgl_features,r_baseline_mgl_features) %>% 
  dplyr::distinct()

@@

# now we count rules, not words. easier.
mgl_rule_count = mgl_baseline_rules_raw %>% # take raw rules
  dplyr::filter( # throw out non-category verbs and non reg/irreg rules (in our definition)
    !is.na(category),
    !is.na(rule.type)
  ) %>% 
  dplyr::mutate( # make a rule column of the form A -> B / C
    rule = paste0(A, ' ->', B, ' / ', Pfeat)
  ) %>% 
  dplyr::distinct(rule,category,rule.type) %>% # keep DISTINCT rules
  dplyr::count(category, rule.type) %>%  # count them
  tidyr::pivot_wider(names_from = rule.type, values_from = n) %>% # stretch table
  dplyr::mutate( # make row sum
    sum = irregular + regular
  ) %>% 
  dplyr::rename( # make nice names
    `regular rule` = regular,
    `irregular rule` = irregular
  )

# glue on column sums
mgl_rule_count = rbind(mgl_rule_count, c('sum', sum(mgl_rule_count$`irregular rule`), sum(mgl_rule_count$`regular rule`), sum(mgl_rule_count$sum)))

@@

<<mgltable, echo=F, eval=T, results=tex>>=
print(xtable::xtable(mgl_table, digits=2, label="mglruleses", caption="The number of unique rules posited by the MGL in each verb class"))

@@

% directly piping in this table from R would be difficult, but it's lookupable
\begin{table}[ht]
\centering
\begin{tabular}{lll}
  \hline
alternation & preceding environment\\ 
  \hline
\textipa{il} $\rightarrow{}$ \textipa{Elt} & obstruent \\ 
\textipa{ip} $\rightarrow{}$ \textipa{Ept} & \{j, l, r, w\} \\ 
\textipa{ip} $\rightarrow{}$ \textipa{Ept} & dorsal consonant\\
\textipa{ip} $\rightarrow{}$ \textipa{Ept} & l \\ 
\textipa{ip} $\rightarrow{}$ \textipa{Ept} & any consonant \\
\textipa{il} $\rightarrow{}$ \textipa{Elt} & fricative or nasal consonant \\
$\emptyset{}$ $\rightarrow{}$ \textipa{d}    & alveolars \\ 
$\emptyset{}$ $\rightarrow{}$ \textipa{d}    & \{b, m\} \\ 
$\emptyset{}$ $\rightarrow{}$ \textipa{d}    & alveolar fricative or nasal consonant \\ 
$\emptyset{}$ $\rightarrow{}$ \textipa{d}    & \{b, m, p\} \\
   \hline
\end{tabular}
\caption{The unique rules posited for the KEPT class}
\label{mglruleseskept}
\end{table}

@@

##########################################
# VIF for GLMM-s, from Florian Jaeger

@@

# while the glm syntax estimates an intercept even if you don't spell it out, I spell it out, because, along with the syntax for the random intercept, it makes what's happening more evident to me.

# all fits use the bobyqa optimiser which tends to have fewer convergence issues in my experience

@@

save(lme.fit1, file = 'models/si_fit1.rda')
save(lme.fit1b, file = 'models/si_fit1b.rda')
save(lme.fit1c, file = 'models/si_fit1c.rda')
vif.mer(lme.fit1)
vif.mer(lme.fit1c)

@@

# we don't report this comparison here or in the main text, but the models are stored for reference

@@

<<main_model3, echo=F,eval=T, results=tex>>=
load('models/si_fit1.rda')
print(xtable::xtable(summary(lme.fit1)$coef, digits = 3, label = 'regressiontable1', caption = 'Experimental Data, Post-test Regression Model Summary. Model formula: post-test regular response ~ verb baseline mean x lexical typicality + participant pretest mean + regularization shift + (1 + verb.baseline.mean | participant) + (1 | verb)'))
@

@@

<<Bayes factor, echo=F, eval=F>>=

# we keep random vs atypical

test2b = test2 %>% 
  dplyr::filter(
    bot_lexical_typicality != 'typical'
  )

@@

# summary(fit_power1)
# summary(fit_power2)
save(fit_power1, file = 'models/si_fit_power1.rda')
save(fit_power2, file = 'models/si_fit_power2.rda')

# a simple way to calculate Bayes factors, should you wish to do so:
exp((BIC(fit_power2) - BIC(fit_power1))/2)
@

@@

# calculating concordance indices.
# we DON'T load Hmisc in itself as it messes up function names

@@

# get correlation between mgl and gcm scores in baseline
mgl_gcm_cor = baseline %>% 
  dplyr::distinct(word,r_baseline_mgl_features,baseline_gcm_features) %>% 
  dplyr::summarise(
    cor = cor(r_baseline_mgl_features,baseline_gcm_features)
  ) %>% 
  dplyr::pull() %>% 
  round(3)

@@

# simple baseline words data frame
baseline_words = baseline %>% 
  dplyr::distinct(word, r_baseline_mgl_features, r_baseline_gcm_features)

# fitting models 1 and 2 on it
celex_m1 = lm(r_baseline_gcm_features ~ 1 + r_baseline_mgl_features, data = baseline_words)
celex_m2 = lm(r_baseline_mgl_features ~ 1 + r_baseline_gcm_features, data = baseline_words)

# getting residuals from models and putting them back in this data frame
baseline_words = baseline_words %>% 
  dplyr::mutate(
    r_baseline_gcm_features_residual = resid(celex_m1),
    r_baseline_mgl_features_residual = resid(celex_m2)
  )

# joining it with entire baseline data
baseline = left_join(baseline, baseline_words, by = c("word", "r_baseline_mgl_features", "r_baseline_gcm_features"))

# fitting models 3 and 4
celex_m3 = glmer(regular ~ 
                1 + r_baseline_mgl_features + r_baseline_gcm_features_residual + 
                (1 + r_baseline_mgl_features + r_baseline_gcm_features_residual | subject), 
                data = baseline, 
                family = binomial ( link = logit ),
                control=glmerControl(optimizer="bobyqa"))

@@

save(celex_m3, file = 'models/si_celex_m3.rda')
save(celex_m4, file = 'models/si_celex_m4.rda')

@@

<<celex_resid1, echo=F,eval=T, results=tex>>=
load('models/si_celex_m3.rda')
print(xtable::xtable(summary(celex_m3)$coef, digits = 3, label = 'resid1', caption = 'MGL scores and GCM residuals, baseline data'))
@

<<celex_resid2, echo=F,eval=T, results=tex>>=
load('models/si_celex_m4.rda')
print(xtable::xtable(summary(celex_m4)$coef, digits = 3, label = 'resid2', caption = 'GCM scores and MGL residuals, baseline data'))
@

@@

\section{Model fits on the post-test data}\label{mglgcmposttest}

@@

# concordance indices of everything on everything, saving into variables, piping into table below

@@

# fitting models 5 and 6

@@

save(indiv_m5, file = 'models/si_indiv_m5.rda')

@@

save(indiv_m6, file = 'models/si_indiv_m6.rda')

@@

load('models/si_indiv_m5.rda')
print(xtable::xtable(summary(indiv_m5)$coef, digits = 3, label = 'indiv1', caption = 'Celex-GCM and individual score difference, post-test data'))

@@

<<indiv_diff2, echo=F,eval=T, results=tex>>=
load('models/si_indiv_m6.rda')
print(xtable::xtable(summary(indiv_m6)$coef, digits = 3, label = 'indiv2', caption = 'Celex-MGL and individual score difference, post-test data'))
@

@@

save(comp_fit1, file = 'models/si_comp_fit1.rda')
save(comp_fit1b, file = 'models/si_comp_fit1b.rda')
save(comp_fit1c, file = 'models/si_comp_fit1c.rda')
save(comp_fit1d, file = 'models/si_comp_fit1d.rda')
save(comp_fit1e, file = 'models/si_comp_fit1e.rda')

@@

# getting a whole lot of chisq values. p values are not reported, but can be guessed from chisq values / are lookupable.

load('models/si_comp_fit1.rda')
load('models/si_comp_fit1b.rda')
load('models/si_comp_fit1c.rda')
load('models/si_comp_fit1d.rda')
load('models/si_comp_fit1e.rda')

@@

<<comparison_model3, echo=F,eval=T, results=tex>>=
print(xtable::xtable(summary(comp_fit1)$coef, digits = 3, label = 'comp3', caption = 'Celex-MGL and Individual-GCM predict variation in the post-test'))
@

@@

# big combination figure.

# first, verb baseline regularisation averages and post-test averages across lex typicality
first_panel = test2 %>% 
  dplyr::group_by(
    word,
    bot_lexical_typicality,
    baseline_av # baseline average is there already
  ) %>%
  dplyr::summarise(
    post_reg_mean = resp_post_reg %>% # we get mean post test regularisation per lex typ
      mean()
  ) %>% 
  ungroup() %>% 
  dplyr::mutate( # we scale everything
    baseline = scales::rescale(baseline_av), # we call them baseline and posttest respectively
    posttest = scales::rescale(post_reg_mean),
    type = 'Participant responses' # we name this
  )

# same for gcm weights. 
second_panel = test2 %>% 
  dplyr::group_by(
    word,
    bot_lexical_typicality,
    r_baseline_gcm_features
  ) %>% 
  dplyr::rename(
    baseline = r_baseline_gcm_features
  ) %>% 
  dplyr::summarise(
    posttest = mean(r_individual_gcm_features)
  ) %>% 
  dplyr::mutate(
    type = 'GCM'
  )

# same for MGL weights
third_panel = test2 %>% 
  dplyr::group_by(
    word,
    bot_lexical_typicality,
    r_baseline_mgl_features
  ) %>% 
  dplyr::rename(
    baseline = r_baseline_mgl_features
  ) %>% 
  dplyr::summarise(
    posttest = mean(r_individual_mgl_features)
  ) %>% 
  dplyr::mutate(
    type = 'MGL'
  )

# bind the three data frames, set factor order for type so pretty plot
three_studies = bind_rows(first_panel,second_panel,third_panel) %>% 
  dplyr::mutate(
    type = factor(type, levels = c('Participant responses', 'GCM', 'MGL'))
  )

# plot this, x is baseline, y is posttest.
# colour AND shape are lexical typicality.
# split panels across raw data / gcm / mgl.
# add loess smooths.
p1 = ggplot(three_studies, aes(x = baseline, y = posttest, colour = bot_lexical_typicality, shape = bot_lexical_typicality, linetype = bot_lexical_typicality)) +
  guides(size = F) +
  geom_point() +
  geom_smooth(se = F, method = 'loess', formula = 'y ~ x') +
  facet_wrap( ~ type) +
  scale_shape_discrete(name="Lexical\ntypicality") +
  # scale_colour_discrete(name="Lexical\ntypicality") +
  xlab('verb baseline regularization') +
  ylab('verb ESP regularization') +
  guides(linetype = F) +
  scale_colour_brewer(palette = 'Set2', name="Lexical\ntypicality")
